Global Distribution of Unintended Pregnancies and Contraceptive Usage

### 2024-04-16 code/analysis-code/statistical-analysis.qmd
Part 5 Updates
Critically analyzed each model and chose the best one. Added a summary of findings with interpretation to the manuscript. Attempted to fix the table and figure captions again.

## Data fitting and statistical analysis 
The outcome of interest in all of the following models is the percentage of unintended pregnancies associated with usage of short acting reversible contraception methods. The predictors used in our models include total number of women using short acting reversible methods standardized by population size, region, percent of women currently married, percent of women never married, total rate of maternal deaths, rate of maternal deaths due to unintended pregnancy, and rate of maternal deaths due to abortion. We are considering adding another predictor using a political freedom index for each country.

@@

## Assessing Model Performance
It is important to consider appropriate metrics to measure model performance. We will use RMSE and MAE at this point to compare 3 models: one with the number of women using short acting reversible contraception predictor, one using the percent never married predictor, and one using the rate of maternal deaths due to abortion predictor. We chose these predictors because they have the strongest associations with the outcomes based on the single linear regression models shown above. We chose these performance metrics because they are commonly used for linear regression models like the ones we have created. RMSE provides an average magnitude of the error in the predictors, and MAE provides an average absolute difference between the predicted and actual values. We include both metrics to increase the validity of our results. 
```{r}
# set seed for reproducibility 
set.seed(2468)

@@

Since we have evidence that our model overfits the data, the last technique we want to use to explore our full model is Lasso regression. This regression analysis shrinks the coefficients towards the mean, which allows it to perform subset selection. Coefficients can reach zero, and that results in that variable being removed from the model. This technique could reveal predictors that should not be included in our final model, which could reduce the risk of overfitting. 
```{r}
# set seed for reproducibility
set.seed(2468)

@@

# save the random forest figure 
figurefile4 <- here("results", "figures", "final-figures", "randomforestpred.png")
ggsave(filename = figurefile4, plot = figure4)
```
The random forest model has an RMSE of 4.361363, an MAE of 3.386193, and an R-squared of 0.8369754. Compared to the previous model with an RMSE of 4.361363, an MAE of 3.386193, and an R-squared of 0.820600, the random forest model performs very similarly. The RMSE and MAE values are exactly the same beause the same 5 predictors are used for the outcome, but the R-squared value shows that the random forest model performs slightly better. Since a higher R-squared value represents smaller differences between the observed and predicted values, we can conclude that this model is predicting our outcome with slightly higher accuracy. The plot of the predicted versus observed values shows strong correlation for predicted values less than 15, but the deviation from the expected values increases as the predicted values increase.

@@

# save plot 
figure_file5 = here("results", "figures", "final-figures", "randomforesttune.png")
ggsave(filename = figure_file5, plot=rf_tune_plot) 
```
Based on the red line that shows the lowest possible RMSE score, our random forest model performs the best with a node size of 1 and 4 predictors. This begs us to ask the question of which one of the 5 predictors included in our model should be dropped to improve model performance because all previous performance metrics suggested that including all 5 predictors created the best model.

Choose final model with justification

>>>>

## Data fitting and statistical analysis 
The outcome of interest in all of the following models is the percentage of unintended pregnancies associated with usage of short acting reversible contraception methods. The predictors used in our models include total number of women using short acting reversible methods standardized by population size, region, percent of women currently married, percent of women never married, total rate of maternal deaths, rate of maternal deaths due to unintended pregnancy, and rate of maternal deaths due to abortion.

@@

## Assessing Model Performance
It is important to consider appropriate metrics to measure model performance. We will use RMSE and MAE at this point to compare 3 models: one with the number of women using short acting reversible contraception predictor, one using the percent currently married predictor, and one using the rate of maternal deaths due to abortion predictor. We chose these predictors because they have the strongest associations with the outcomes based on the single linear regression models shown above. We chose these performance metrics because they are commonly used for linear regression models like the ones we have created. RMSE provides an average magnitude of the error in the predictors, and MAE provides an average absolute difference between the predicted and actual values. We include both metrics to increase the validity of our results. 
```{r}
# set seed for reproducibility 
set.seed(2468)

@@

Since we have evidence that our model overfits the data, we want to explore our full model using Lasso regression. This regression analysis shrinks the coefficients towards the mean, which allows it to perform subset selection. Coefficients can reach zero, and that results in that variable being removed from the model. This technique could reveal predictors that should not be included in our final model, which could reduce the risk of overfitting. 
```{r}
# set seed for reproducibility
set.seed(2468)

@@

# save the random forest figure 
figurefile4 <- here("results", "figures", "final-figures", "randomforestpred.png")
ggsave(filename = figurefile4, plot = figure4)
```
The random forest model has an RMSE of 4.361363, an MAE of 3.386193, and an R-squared of 0.8369754. Compared to the previous model with an RMSE of 4.361363, an MAE of 3.386193, and an R-squared of 0.820600, the random forest model performs very similarly. The RMSE and MAE values are exactly the same because the same 5 predictors are used for the outcome, but the R-squared value shows that the random forest model performs slightly better. Since a higher R-squared value represents smaller differences between the observed and predicted values, we can conclude that this model is predicting our outcome with slightly higher accuracy. The plot of the predicted versus observed values shows strong correlation for predicted values less than 15, but the deviation from the expected values increases as the predicted values increase.

@@

# save plot 
figure_file5 = here("results", "figures", "final-figures", "randomforesttune.png")
ggsave(filename = figure_file5, plot=rf_tune_plot)

```
Based on the red line that shows the lowest possible RMSE score, our random forest model performs the best with a node size of 1 and 4 predictors. This begs us to ask the question of which one of the 5 predictors included in our model should be dropped to improve model performance because all previous performance metrics suggested that including all 5 predictors created the best model.

## Choosing a final model 
After considering each of the 10 models we created throughout the statistical analysis process, the random forest model with 5 predictors is the best overall choice for predicting our outcome of percentage of unintended pregnancies due to short acting reversible methods. The collinearity between the total rate of maternal deaths, rate of maternal deaths due to abortion, and rate of maternal deaths due to unintended pregnancies justified why they should not be included in the same model. We chose to remove the total maternal mortality predictor to avoid the issue of collinearity because the other two predictors in this category had lower RMSE and MAE performance metrics. The same reasoning explains why the percent of currently married women and the percent of never married women predictors should not be included in the same model. We chose to include the percent of currently married women in the final model because it had the strongest relationship with the outcome. The four most promising models were the single predictor models with population size using short acting reversible contraception, percent of currently married women, and rate of maternal deaths due to abortion, respectively, along with the full model with all five predictors. The single model with the region variable only performed poorly with an RMSE of 18.11857, so it was not considered in the final running for the best model. The models with percent of currently married women and rate of maternal deaths due to abortion had comparable RMSE scores of 10.175909 and 10.141275, respectively. Large improvement in performance metrics was observed when considering the model with population size using short acting reversible contraception based on an RMSE of 5.143319. The full model with all five predictors, including short acting reversible contraception usage, region, percent of currently married women, rate of maternal mortality due to abortion, and rate of maternal mortality due to unintended pregnancies, performed the best with an RMSE of 4.361363. We felt it was important to include the region predictor in the final model even though it did not perform well in a model on its own because of the large differences in the coefficients for the region predictor in a linear model with the outcome. Latin America/the Caribbean and Asia have much larger effects on the outcome than Europe and Oceania, so we argue for the inclusion of this variable in the final model to capture this information.

Once we decided that the full model with 5 predictors was the best choice, we used cross-validation with LASSO regression and random forest modeling to explore subset selection. The LASSO regression kept all 5 predictors, and the coefficients for the Asia, Europe, and Latin America/the Caribbean predictors were the largest behind the sarc_standard predictor. The smallest coefficient was assigned to the percent of currently married women predictor. The parameters were tuned for the LASSO regression, which increased our confidence in using a penalty of 0.1. The random forest model generated the same RMSE of 4.361363 as the LASSO regression model, but it has a higher R-squared value of 0.8369754. Since the random forest model explains 83.69% of the variation between the predicted and observed outcomes compared to 82.06% explained by the LASSO regression model, it is slightly better at predicting the outcome. After tuning the parameters for this model, the number of randomly selected predictors that produces the lowest RMSE is 4, not 5. While we cannot be sure which predictor the model is suggesting to remove, we hypothesize that it is the percent of currently married women based on the information from the LASSO regression. However, the improvement in RMSE when using 4 predictors instead of 5 is minimal, and the minimal node size variable is much more impactful on the RMSE of the model. Based on all of this information, we suggest using the full model with all 5 predictors with a LASSO regression because it accurately portrays the importance of each predictor while providing a penalty to predictors that do not meaningfully contribute to the model. While this model has a slightly lower R-squared value compared to the random forest model, the evidence of overfitting is so strong that this difference is negligible. Cross-validation of the model with all 5 predictors showed that predictions from testing data deviated from the expected relationship with larger variance than the predictions from the training data. Since the model did not perform well when used on test data, we conclude that there is overfitting in our model. Unfortunately, it is likely that any of these models are not optimal choices for predicting our outcome based on the poor performance of our model on testing data. While we suggest using the LASSO regression with all 5 predictors to predict our outcome because it is slightly easier to interpret, future researchers who are willing to sacrifice interpretation for predictive power should use the random forest model with 4 predictors based on the slightly higher predictive performance shown by the R-squared performance metric.

### 2024-04-16 products/manuscript/Manuscript.qmd
Part 5 Updates
Critically analyzed each model and chose the best one. Added a summary of findings with interpretation to the manuscript. Attempted to fix the table and figure captions again.

The exploratory data analysis revealed a less skewed distribution of unintended pregnancies among SARC users compared to other contraceptive methods. This led researchers to focus on SARC use for the remainder of the study. The Latin America/Caribbean region displayed the strongest association with unintended pregnancy rates, followed by Asia. The Oceania and Europe region had weaker associations. Marriage status (both currently and never married) was inversely associated with unintended pregnancy rates. This means that as the percent of women who are currently married or never married increases, the unintended preganacy rate decreases. Maternal mortality rates showed a positive correlation with the outcoe.

@@

Overall, this study highlights the influence of regional variations, SARC use, and marital status on unintended pregnancy rates in low- and middle-income countries. The findings suggest a need for further investigation into the factors influencing maternal mortality rates and unintended pregnancies.

@@

Step four is to critically analyze the performance of our models to determine which one performs the best in predicting our outcome, which is also done in the statistical-analysis.qmd file in the code > analysis-code folder. We chose 2 performance metrics to find for each of our 9 models: Root Mean Square Error (RMSE) and Mean Absolute Error (MAE). We chose these performance metrics because they are commonly used for linear regression models like the ones we have created. RMSE provides an average magnitude of the error in the predictors, and MAE provides an average absolute difference between the predicted and actual values. We include both metrics to increase the validity of our results. Once we decided on the final model, we use cross-validation to determine how well it will perform on data it has not seen before. This involves splitting the data into testing and training pieces , which allows us to compare the predictions created using both datasets. To examine the uncertainty in the model, we will use the bootstrapping method to resample our data 100 times to find confidence intervals for our predictions. This process is similar to cross-validation because it relies on resamling the training data. Lastly, we used a LASSO regression to add a penalty to each parameter included in the model to determine if any variables should be removed to improve model performance. It is important to tune the hyperparameters associated with LASSO regression to determine which model fits best. We use a grid search to ensure that we try all combinations of the parameters in our grid. We complete the tuning process with a 5-fold cross-validation repeated 5 times. We added a random forest model to attempt to improve model performance. We used another parameter tuning grid to allow for tuning of 2 variables: `mtry`, the number of variable to possibly split at in each node, and `min_n`, the minimal node size. We set the number of trees to 300 because it strikes a good balance between achieving the best performance metrics and the processing time required to do so. Each of these processes are crucial in determining which model performs best for our dataset. Result tables for each model and test are stored in the results > tables folder. Figures created from cross-validation, boostrapping, LASSO regression, and random forest modeling are stored in the results > figure > final-figures folder. Lastly, we examined the AIC scores for the two best performing models since there is strong evidence for overfitting. Even though cross-validation is a better approach for answering the question of how well a model predicted values in unseen data because it actually resamples the data, the concern of overfitting justifies finding the AIC value in this case. This measure still evaluates the trade-off between good fit to the data, shown by a low cost function, and model complexity based on the number of parameters.

@@

After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen @table1.
```{r}
#| message: false
#| label: table1
#| tbl-cap: "Table 1: Summary Table Of Numeric Variables"
#| echo: FALSE
predictortable<- readRDS(here("results", "tables", "predictortable.rds"))
predictortable
```

We created a multitude of exploratory plots to understand our data better. The first exploratory result shown in @figure1 displays the distribution of the outcome variable: percent of unintended pregnancies due to short acting reversible methods. The `pct_upreg_sarc` variable is slightly skewed to the right with an average around 12%.
```{r}
#| message: false
#| label: figure1
#| fig-cap: "Plot 1: Histogram Depicting Number of Countries With Percent Unintended Pregnancy Using SARC"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_distribution.png"))
```

@@

Lastly, we explored the predictor variables that will be used in our models, and the graphs are shown in @figure5 - @figure10. The Africa region accounts for most of the observations with about 53 countries, followed by Asia that includes. about 35 countries. Latin America/the Caribbean accounts for about 25 countries. Europe and Oceania account for the least number of observations. The distribution of women currently married is approximately normal with an average around 57% and no major outliers. The distribution of women never married is slightly skewed to the right with an average around 30%, but there are a few outliers greater than 60%. The distribution of the total maternal mortality rate is highly skewed to the right with a couple outliers. We would have expected the highest rates to belong to India and China because they have the most people, but it is Chad, Sierra Leone, and South Sudan that account for the highest rates of total maternal mortality. The distribution of the maternal mortality rate from unintended pregnancies is highly skewed to the right with a couple outliers, which is to be expected based on the trend shown in the previous histogram. Chad and South Sudan also account for the highest rates of maternal mortality in this category. The distribution of maternal mortality rate from abortions is also highly skewed to the right. The outliers are accounted for by Gambia and Mauritania in this predictor.
```{r}
#| message: false
#| label: figure5-figure10
#| fig-cap: "Predictor Variable Distributions"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "region_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "currmarried_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "nevmarried_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeaths_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeathsupreg_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeathsabs_distribution.png"))
```

@@

We created one model with all 5 predictors using multiple linear regression, shown in @table4. This model shows the strongest predictor is the number of women using short acting reversible contraception. The next strongest predictor is the European region followed by the Latin America/the Caribbean region and the Asia region, which is a shift from the model with the singular region predictor. The Oceania region now has a negative association with the outcome. The currently married predictor is associated with a weak decrease in the outcome. The coefficient estimates for the rate of maternal death due to unintended pregnancies and abortion both show a very weak positive association with the outcome. 
```{r}
#| message: false
#| label: table4
#| fig-cap: "Table 4: Multiple Linear Regression with 5 Predictors"
#| echo: FALSE
readRDS(here("results","tables", "resulttable8.rds")) 
```
To determine which model performed the best, we started by examining the RMSE and MAE performance metrics. We found that the model with the standardized total count of women using the short acting reversible contraception method performed the best in a single predictor model compared to the percent of currently married women and the rate of maternal mortality due to abortion. @table5 shows the RMSE and MAE values for each of these three models compared to the null model with no predictors. The model with the population size using SARCs is the only single predictor model that performs better than the null. The full model with all 5 predictors peforms the best with the lowest RMSE and MAE values. 
```{r}
#| message: false
#| label: table5
#| fig-cap: "Table 5: Performance Metric Comparisons"
#| echo: FALSE
readRDS(here("results","tables", "resulttable14.rds")) 
```
We used cross-validation of the two best models to determine how well the model performs on data it has not seen yet by splitting the data into testing and training pieces. The cross-validated results, shown in @table6 lead to the same conclusion: the multiple linear regression model with 5 predictors performs better. The RMSE value for the single predictor model with `sarc_standard` is 4.822226, which is lower the the first RMSE value of 5.143319. While the RMSE value for the multiple linear regression model is slightly higher than the first time it was calculated, it is still lower than the single predictor model. Additionally, the standard error of the RMSE for the multiple linear regression model is 0.2768, which is lower than the standard error of the RMSE for the single predictor model of 0.3528. This means there is higher variance in the single predictor model. The R-squared value of 0.8206 for the multiple linear regression model is higher than 0.7861 for the single predictor model, which means the multiple linear regression model accounts for more of the variability in the data. Overall, the multiple linear regression model performs better.
```{r}
#| message: false
#| label: table6
#| fig-cap: "Table 6: Cross Validation Performance Metric Comparisons"
#| echo: FALSE
readRDS(here("results","tables", "resulttable15.rds")) 
```

@@

@@

## Summary and Interpretation
_Summarize what you did, what you found and what it means._
critically think about each model (same number of predictors so we recommend random forest) all not great

>>>>

The exploratory data analysis revealed a less skewed distribution of unintended pregnancies among SARC users compared to other contraceptive methods. This led researchers to focus on SARC use for the remainder of the study. The Latin America/Caribbean region displayed the strongest association with unintended pregnancy rates, followed by Asia. The Oceania and Europe region had weaker associations. Marriage status, both currently and never married, was inversely associated with unintended pregnancy rates. This means that as the percent of women who are currently married or never married increases, the unintended pregnancy rate decreases. Maternal mortality rates showed a positive correlation with the outcome.

@@

Overall, this study highlights the influence of regional variations, short acting reversible contraception use, and marital status on unintended pregnancy rates in low- and middle-income countries. The findings suggest a need for further investigation into the factors influencing maternal mortality rates and unintended pregnancies.

@@

Step four is to critically analyze the performance of our models to determine which one performs the best in predicting our outcome, which is also done in the statistical-analysis.qmd file in the code > analysis-code folder. We chose 2 performance metrics to find for each of our 9 models: Root Mean Square Error (RMSE) and Mean Absolute Error (MAE). We chose these performance metrics because they are commonly used for linear regression models like the ones we have created. RMSE provides an average magnitude of the error in the predictors, and MAE provides an average absolute difference between the predicted and actual values. We include both metrics to increase the validity of our results. Once we decided on the final model, we use cross-validation to determine how well it will perform on data it has not seen before. This involves splitting the data into testing and training pieces , which allows us to compare the predictions created using both datasets. To examine the uncertainty in the model, we will use the bootstrapping method to resample our data 100 times to find confidence intervals for our predictions. This process is similar to cross-validation because it relies on resamling the training data. Lastly, we used a LASSO regression to add a penalty to each parameter included in the model to determine if any variables should be removed to improve model performance. It is important to tune the hyperparameters associated with LASSO regression to determine which model fits best. We use a grid search to ensure that we try all combinations of the parameters in our grid. We complete the tuning process with a 5-fold cross-validation repeated 5 times. We added a random forest model to attempt to improve model performance. We used another parameter tuning grid to allow for tuning of 2 variables: `mtry`, the number of variable to possibly split at in each node, and `min_n`, the minimal node size. We set the number of trees to 300 because it strikes a good balance between achieving the best performance metrics and the processing time required to do so. Each of these processes are crucial in determining which model performs best for our dataset. Result tables for each model and test are stored in the results > tables folder. Figures created from cross-validation, boostrapping, LASSO regression, and random forest modeling are stored in the results > figure > final-figures folder.

@@

After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen in @tbl-1.
```{r}
#| message: false
#| label: tbl-1
#| tbl-cap: "Table 1: Summary Table Of Numeric Variables"
#| echo: FALSE
predictortable<- readRDS(here("results", "tables", "predictortable.rds"))
predictortable
```

We created a multitude of exploratory plots to understand our data better. The first exploratory result shown in @fig-1 displays the distribution of the outcome variable: percent of unintended pregnancies due to short acting reversible methods. The `pct_upreg_sarc` variable is slightly skewed to the right with an average around 12%.
```{r}
#| message: false
#| label: fig-1
#| fig-cap: "Plot 1: Histogram Depicting Number of Countries With Percent Unintended Pregnancy Using SARC"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_distribution.png"))
```

@@

Lastly, we explored the predictor variables that will be used in our models, and the graphs are shown in @figure5 - @figure10. The Africa region accounts for most of the observations with about 53 countries, followed by Asia that includes about 35 countries. Latin America/the Caribbean accounts for about 25 countries. Europe and Oceania account for the least number of observations. The distribution of women currently married is approximately normal with an average around 57% and no major outliers. The distribution of women never married is slightly skewed to the right with an average around 30%, but there are a few outliers greater than 60%. The distribution of the total maternal mortality rate is highly skewed to the right with a couple outliers. We would have expected the highest rates to belong to India and China because they have the most people, but it is Chad, Sierra Leone, and South Sudan that account for the highest rates of total maternal mortality, which each belong to the African region. The distribution of the maternal mortality rate from unintended pregnancies is highly skewed to the right with a couple outliers, which is to be expected based on the trend shown in the previous histogram. Chad and South Sudan also account for the highest rates of maternal mortality in this category. The distribution of maternal mortality rate from abortions is also highly skewed to the right. The outliers are accounted for by Gambia and Mauritania in this predictor, which both belong to the African region.
```{r}
#| message: false
#| label: figure5-figure10
#| fig-cap: "Predictor Variable Distributions"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "region_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "currmarried_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "nevmarried_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeaths_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeathsupreg_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeathsabs_distribution.png"))

```

@@

We created one model with all 5 predictors using multiple linear regression, shown in @table4. This model shows the strongest predictor is the number of women using short acting reversible contraception. The next strongest predictor is the European region followed by the Latin America/the Caribbean region and the Asia region, which is a shift from the model with the singular region predictor. The Oceania region now has a negative association with the outcome. The currently married predictor is associated with a weak decrease in the outcome. The coefficient estimates for the rate of maternal death due to unintended pregnancies and abortion both show a very weak positive association with the outcome. 
```{r}
#| message: false
#| label: table4
#| fig-cap: "Table 4: Multiple Linear Regression with 5 Predictors"
#| echo: FALSE
readRDS(here("results","tables", "resulttable8.rds")) 
```
To determine which model performed the best, we started by examining the RMSE and MAE performance metrics. We found that the model with the standardized total count of women using the short acting reversible contraception method performed the best in a single predictor model compared to the percent of currently married women and the rate of maternal mortality due to abortion. @table5 shows the RMSE and MAE values for each of these three models compared to the null model with no predictors. The model with the population size using SARCs is the only single predictor model that performs better than the null. The full model with all 5 predictors performs the best with the lowest RMSE and MAE values. 
```{r}
#| message: false
#| label: table5
#| fig-cap: "Table 5: Performance Metric Comparisons"
#| echo: FALSE
readRDS(here("results","tables", "resulttable14.rds")) 
```
We used cross-validation of the two best models to determine how well the model performs on data it has not seen yet by splitting the data into testing and training pieces. The cross-validated results, shown in @table6 lead to the same conclusion: the multiple linear regression model with 5 predictors performs better. The RMSE value for the single predictor model with `sarc_standard` is 4.822226, which is lower the the first RMSE value of 5.143319. While the RMSE value for the multiple linear regression model is slightly higher than the first time it was calculated, it is still lower than the single predictor model. Additionally, the standard error of the RMSE for the multiple linear regression model is 0.2768, which is lower than the standard error of the RMSE for the single predictor model of 0.3528. This means there is higher variance in the single predictor model. The R-squared value of 0.8206 for the multiple linear regression model is higher than 0.7861 for the single predictor model, which means the multiple linear regression model accounts for more of the variability in the data. Overall, the multiple linear regression model performs better.
```{r}
#| message: false
#| label: table6
#| fig-cap: "Table 6: Cross Validation Performance Metric Comparisons"
#| echo: FALSE
readRDS(here("results","tables", "resulttable15.rds")) 
```

@@

As seen in the statistical analysis file, we created 10 different model options with unique combinations of our 7 predictors. After considering effect sizes and performance metrics for each model, we have decided that the LASSO regression model with 5 predictors is the best overall choice for predicting our outcome of percentage of unintended pregnancies due to short acting reversible methods. After considering each of the 10 models we created throughout the statistical analysis process, the random forest model with 5 predictors is the best overall choice for predicting our outcome of percentage of unintended pregnancies due to short acting reversible methods. We used cross-validation with LASSO regression and random forest modeling to explore subset selection of our full model with all 5 predictors. The LASSO regression kept all 5 predictors, and the coefficients for the Asia, Europe, and Latin America/the Caribbean predictors were the largest behind the sarc_standard predictor. The smallest coefficient was assigned to the percent of currently married women predictor. We suggest using the full model with all 5 predictors with a LASSO regression because it accurately portrays the importance of each predictor while providing a penalty to predictors that do not meaningfully contribute to the model. While this model has a slightly lower R-squared value compared to the random forest model, the evidence of overfitting is so strong that this difference is negligible. Cross-validation of the model with all 5 predictors showed that predictions from testing data deviated from the expected relationship with larger variance than the predictions from the training data. Since the model did not perform well when used on test data, we conclude that there is overfitting in our model. Unfortunately, it is likely that any of these models are not optimal choices for predicting our outcome based on the poor performance of our model on testing data. While we suggest using the LASSO regression with all 5 predictors to predict our outcome because it is slightly easier to interpret, future researchers who are willing to sacrifice interpretation for predictive power should use the random forest model with 4 predictors based on the slightly higher predictive performance shown by the R-squared performance metric.

@@

## Summary and Interpretation
Our exploratory data analysis allowed us to examine the distribution of our outcome of interest, percentage of unintended pregnancies due to short acting reversible contraception, and the predictors used to describe the outcome. Most low- and middle-income countries were from the African region, followed by Asia and Latin America/the Caribbean. The Latin America/the Caribbean region experienced the highest percentage of unintended pregnancies followed by Europe and Asia. This association is likely explained by the fact that women in these regions have better access to short acting reversible contraception methods because we would have expected the African region to experience the highest rate of our outcome based on the number of countries included from that region. After standardizing the number of women who use short acting reversible contraception based on total population size, it is clear the women in the Latin America/the Caribbean region use this contraception method at the highest rates, followed my the Asian region. Further analysis of the region predictor in a linear model with the outcome confirmed this finding by showing that Latin America/the Caribbean has the strongest association with the outcome, followed by Asia. Asia accounted for the highest percentage of currently married women. The highest rate of maternal mortality due to abortion and highest rate of maternal mortality due to unintended pregnancies are both accounted for by the African region. This cannot be explained by having the highest number of reproductive age women because the Asian region has 6 times the number of reproductive age women compared to the African region. This discrepancy is likely due to poorer access to maternal healthcare in Africa compared to Asia.

Our full statistical analysis of the final model with 5 predictors included the standardized number of women using short acting reversible methods, region, percent of currently married women, rate of maternal mortality due to abortion, and rate of maternal mortality due to unintended pregnancies as predictors. Cross-validation using LASSO regression and random forest modeling techniques did not result in an increase in performance metrics, and the model is strongly overfitting the data based on the predicted versus observed outcomes between the testing and training data. Based on the model using LASSO regression, the number of women using short acting reversible contraception is the strongest predictor of the outcome, which is to be expected because women must be using this form of contraception to have an unintended pregnancy due to failure of this contraceptive method in the first place. The most influential factor in this model is the region variable with European, Asian, and Latin American regions having the strongest effects on the outcome. The percent of currently married women, rate of maternal mortality due to abortion, and rate of maternal mortality due to unintended pregnancies as predictors are all negatively associated with the outcome in this model. This implies that the percentage of unintended pregnancies decreases when a higher percentage of women are currently married and when the country experiences a higher maternal mortality rate based on abortion and unintended pregnancies. However, these effects are so small that it is likely these predictors do not have a real effect on the outcome. This analysis was successful in determining that region does have a significant effect on the percentage of unintended pregnancies due to short acting reversible methods with the Latin America/the Caribbean region experiencing the outcome at the highest rates.
